### 2009-04-18 tex/response.txt
sommboost: new version

Dear Editor,

Please find below the response to the referees' comments for the
submission "Substructure boosts to dark matter annihilation from
Sommerfeld enhancement" by myself (submission # DC10513).

I have made changes following the referee's suggestions and have also
update figures 1, 2, 5, 6, and 7 to make the label in these plots more
clear to the reader.

A detailed list of the main changes is:

- I slightly rewrote the paragraph containing Eq. 20 to include the
definition of the virial radius;

- I fixed a type which the referee found in the paragraph below Eq. 39;

- I changed the color of the label in figures 1, 2, 5, 6, and 7 from
  black to white;

- I enlarged figures 9, 10, 12, 14, arranging the two panels now below
  each other. I updated the text and the caption of figure 10 to
  reflect this change.

- In the conclusion I elaborated a fit further on the implications of
  the possibility that the Galactic halo consists for a large part of
  cold dark matter streams, adding a reference to a paper by Afshordi,
  Mohayaee, and Bertschinger which discusses boosts due to such tidal
  debris.

- I also updated all of the references to papers which have gotton
  published since I first submitted this paper.

In addition to these I made a number of cosmetic changes and slight
rewordings to the text to improve the clarity of the exposition, the
most important of which are:

- Some rewordings in the abstract (including spelling out the proper
  name of the Fermi telescope, and emphasizing more clearly what is
  new in the present paper);

- I renamed Sec. III A "Some details of the particle physics model"
  instead of "Some details of the particle physics"

- I moved up the sentence in the paragraph above Eq. 13 which states
  that the DM annihilation flux depends on the density squared to the
  beginning of that section, such that the point of discussing the
  dark matter profile at length is clear at the start of the section
  on the shape of the profile.

Please contact me by the information given below with any questions
regarding this re-submission.

With kind regards,

Jo Bovy

-
Jo Bovy
Ctr. for Cosmology & Particle Physics, 
Dept. of Physics
New York University
4 Washington Place
New York, NY 10003
jb2777@nyu.edu

In what follows the referee's comments are indented.

==================================================================

	Dear Dr. Bovy:

	The manuscript ``Substructure boosts to dark matter
	annihilation from Sommerfeld enhancement'' (dc10513) by Bovy,J
	has been reviewed by one of our referees.  Comments from the
	report are enclosed for your consideration.

	When you resubmit your manuscript, please include a summary of
	the changes made, and a brief response to all recommendations
	and criticisms.

	Sincerely,

	Rashmi Ray
        Senior Assistant Editor
        Physical Review D

	----------------------------------------------------------------
	Report:
	----------------------------------------------------------------
	This is a very good paper. It makes a careful, intelligent
	contribution to the hot topic of boost factors in Galactic
	dark matter annihilation.  I am also impressed with the
	clarity of exposition.

	I have two small suggestions and a typo to convey to the
	author, for his/her optional consideration.

	1. Typo: toward the end of the paragraph following Eq 39, an
           enhancement of 10^-2 is reported; I believe the author
           means 10^+2.

That is correct, I did mean 10^+2. This has been corrected in the text.

	2. In Eq 20, the virial radius r_vir is used; I don't
	   believe it is defined in the text.

I have added the definition of the virial radius in the paragraph
containing Eq.20 before the definition of the concentration. I have
also moved up the definition of the virial mass from where it was
before (it used to be at the end of that same paragraph).

	3. Figs 9, 10, 12, 14 are too small in the copy I received
	   from APS.  Perhaps this is not the author's issue.

This was my fault. I have replaced these figures with larger versions.
-------------------------------------------------------------------------